Generating Images from Captions with Attention

### 2015-10-09 ram_workshop/ram-paper.tex
update
a

\maketitle
\vspace*{-0.4cm}
\begin{abstract}
\vspace*{-0.15cm}
Motivated by the recent progress in generative models, we introduce a model that generates images from natural language descriptions. The proposed model iteratively draws patches on a canvas, while attending to the relevant words in the description. After training on MSCOCO, we compare our models with several baseline generative models on image generation and retrieval tasks. We then demonstrate our model produces higher quality samples than other approaches and generates images with novel scene compositions corresponding to previously unseen captions in the dataset.   
\end{abstract}

@@

Our proposed model 
defines a generative process of images conditioned on the caption. In particular, captions are represented as a sequence of consecutive words and images are represented as a sequence of patches drawn on canvas over time $t=1,...,T$. Our model can be viewed as utilizing the  sequence-to-sequence framework \citep{ilya_mt}.

@@

\vspace{-0.1in}
\subsection{Image Modelling: the Conditional alignDRAW Network}
\comm{
The DRAW network \cite{gregor_draw} is a sequential probabilistic model generating images by accumulating the output at each iterative step. While the original DRAW network assumes the latent variables are independent, it has shown in \citep{bachman_sdm} the model performance is improved by including the dependencies of latent variables. 
}
To generate images conditioned on the caption information, we extended the DRAW network \citep{gregor_draw} to include caption representation $\hlangall$ at each step, shown in Figure~\ref{fig:figmodel}. The conditional DRAW network is a stochastic recurrent neural network that consists of a set of latent variables $\Lat_t$ at each time step. Unlike the original DRAW network where latent variables are independent unit Gaussians $\mathcal{N}(0, I)$, the latent variables in the proposed alignDRAW model have their mean and variance depend on the previous recurrent hidden states $\hdec_{t-1}$ as in \cite{bachman_sdm}. Formally, the image is generated by iteratively computing the following equations for $t=1,...,T$
\begin{align}
\vspace{-0.1in}
\label{eq:x_hat}
\lat_t &\sim \prior(\Lat_t|\Lat_{1:t-1}) = \mathcal{N}(\mu_{t}(\hdec_{t-1}), \sigma_{t}(\hdec_{t-1})),\\
\hdec_t &= \decoder(\hdec_{t-1}, z_t, s_{t-1}), \\
s_{t} = align(\hdec_{t-1}, & \hlangall); \quad \quad
\canv_t = \canv_{t-1} + \writeop(\hdec_t), 
\label{eq:write}
%\canv_t &= \canv_{t-1} + \writeop(\hdec_t)
\end{align}
where $\writeop$ and $\readop$ are the same attention operators as in \citep{gregor_draw}.
The $align$ function is used to compute the alignment between the input caption and intermediate image generative steps \citep{bahdanau_mt}. Given the caption representation from the language model, $\hlangall = [\hlang_{1}, \hlang_{2}, ..., \hlang_{N}]$, the $align$ operator outputs a dynamic sentence representation $s_t$ at each step through a weighted sum using alignment probabilities $\alpha_{1...N}$:
%\begin{align}
%\vspace{-0.1in}
$s_t=align(\hdec_{t-1}, \hlangall) = \alpha_{1}\hlang_{1} + \alpha_{2}\hlang_{2} + ... + \alpha_{N}\hlang_{N}.$
%\end{align}
The corresponding alignment probabilities $\alpha_{1...N}$ at each step are obtained using the caption representation $\hlangall$ and the hidden state of the generative model $\hdec_{t}$, as in \citep{bahdanau_mt}.
\comm{
\begin{align}
e_{tj} &= v^{T}tanh(U\hlang_{j} + W\hdec_{t} + b)\\
\alpha_{j} &= \frac{exp(e_{tj})}{\sum_{j=1}^{N}exp(e_{tj})}.
\end{align}
}
\comm{
Here $\hlang_{0}$ is initialized to the learned bias.
Setting $\alpha_{1...N}$ to $\frac{1}{N}$ turns the encoder into the vanilla model introduced in \citep{cho_mt} without the attention. 
}

\vspace{-0.2cm}
\subsection{Learning and Generation}

\begin{figure}[!h]
\vspace{-0.3in}
\captionsetup[subfigure]{labelformat=empty}
\begin{center}
\includegraphics[width=0.8\textwidth]{figures/alignDrawAnnotated.pdf}\quad
%
\end{center}
\vspace{-0.2in}
\caption{alignDRAW: Generative model of images conditioned on captions.}
\label{fig:figmodel}
\vspace{-0.2in}
\end{figure}
%\vspace{-0.1in}

Our conditional alignDRAW model is trained to maximize the variational lower bound on the log-likelihood, $\log \sum_{\Lat_{1:T}} \prior(\Lat_{1:T})p(\oimage\given\icaption, \Lat_{1:T})$. The posterior inference is approximated by an inference RNN $\post(\Lat_{1:T}\given\icaption,\oimage)$ shown in the red dashed box in Figure (\ref{fig:figmodel}).   
Overall, the variational objective $\loss$ is defined with the model parameters vector $\theta$ as follows:
\begin{align}
\vspace{-0.1in}
\small
\loss_\theta =  &\expectation_{\post(\Lat_{1:t}\given\icaption,\oimage)}\left[ - \log p(\oimage\given\icaption,\Lat_{1:T}) + \sum_{t=2}^T\kldiv\left(\post(\Lat_t\given\Lat_{1:t-1},\icaption,\oimage)\klBars\prior(\Lat_t\given\Lat_{1:t-1},\icaption)\right) \right] \nonumber \\& + \kldiv\left(\post(\Lat_1\given\oimage)\klBars\prior(\Lat_1\given\icaption)\right).
\end{align}

@@

\vspace{-0.2cm}
\section{Experiments on MS COCO dataset}

In the following subsections, we analyze both the qualitative and quantitative aspects of our model as well as compare its performance with that of other, related generative models\footnote{To see more generated images, go to \url{http://www.cs.toronto.edu/~emansim/cap2im.html}}.
%\vspace{-0.1cm}
%\subsection{Analysis of Generated Images}
First, we wanted to see whether the model understood one of the most basic properties of any object, the color. In Figure~\ref{fig:genimages1}, we generated images of school buses with four different colors: yellow, red, green and blue. Although, there are images of buses with different colors in the training set, all mentioned school buses are specifically colored yellow. Despite that, the model managed to generate images of an object that is visually reminiscent of a school bus that is painted with the specified color.

% \vspace{-0.5cm}
\begin{figure}[!h]
 \vspace{-0.5cm}
\captionsetup[subfigure]{labelformat=empty}
\begin{center}
\subfloat[A \underline{yellow} school bus parked in a parking lot.
]{\includegraphics[width=0.23\textwidth]{figures/new/a-yellow-school-bus-parked-in-a-parking-lot-sharp.png}}\quad
%
\subfloat[A \underline{red} school bus parked in a parking lot.
]{\includegraphics[width=0.23\textwidth]{figures/new/a-red-school-bus-parked-in-a-parking-lot-sharp.png}}\quad
%
\subfloat[A \underline{green} school bus parked in a parking lot.
]{\includegraphics[width=0.23\textwidth]{figures/new/a-green-school-bus-parked-in-a-parking-lot-sharp.png}}\quad
%
\subfloat[A \underline{blue} school bus parked in a parking lot.
]{\includegraphics[width=0.23\textwidth]{figures/new/a-blue-school-bus-parked-in-a-parking-lot-sharp.png}}\quad
%
\end{center}
\vspace{-0.3cm}
\caption{Examples of changing the color while keeping the caption fixed.}
\label{fig:genimages1}
\vspace{-0.3cm}
\end{figure}
Apart from changing the colors of objects, we experimented with changing the background of the scene described in a caption to see whether this would result in the appropriate changes in the generated samples. The task of changing the background of an image is somewhat harder than just changing the color of an object because the model will have to make alterations over a wider visual area. Nevertheless, as shown in Figure~\ref{fig:genimages2}, changing the skies from blue to rainy in a caption as well as changing the grass type from dry to green in another caption resulted in the appropriate changes in the generated image. The nearest images from the training set also indicate that the model was not simply copying the patterns it observed during the learning phase.

\vspace{-0.5cm}
\begin{figure}[!h]
\captionsetup[subfigure]{labelformat=empty}
\begin{center}
\subfloat[]{\includegraphics[width=0.23\textwidth]{figures/new/a-very-large-commercial-plane-flying-in-blue-skies-closest.png}}\quad
%
\subfloat[]{\includegraphics[width=0.23\textwidth]{figures/new/a-very-large-commercial-plane-flying-in-rainy-skies-closest.png}}\quad
%
\subfloat[]{\includegraphics[width=0.23\textwidth]{figures/new/a-herd-of-elephants-walking-across-a-dry-grass-field-closest.png}}\quad
%
\subfloat[]{\includegraphics[width=0.23\textwidth]{figures/new/a-herd-of-elephants-walking-across-a-green-grass-field-closest.png}}\\
\vspace{-0.45cm}
%
\subfloat[A very large commercial plane flying in \underline{blue} skies.
]{\includegraphics[width=0.23\textwidth]{figures/new/a-very-large-commercial-plane-flying-in-blue-skies-sharp.png}}\quad
%
\subfloat[A very large commercial plane flying in \underline{rainy} skies.
]{\includegraphics[width=0.23\textwidth]{figures/new/a-very-large-commercial-plane-flying-in-rainy-skies-sharp.png}}\quad
%
\subfloat[A herd of elephants walking across a \underline{dry} grass field.
]{\includegraphics[width=0.23\textwidth]{figures/new/a-herd-of-elephants-walking-across-a-dry-grass-field-sharp.png}}\quad
%
\subfloat[A herd of elephants walking across a \underline{green} grass field.
]{\includegraphics[width=0.23\textwidth]{figures/new/a-herd-of-elephants-walking-across-a-green-grass-field-sharp.png}}\quad
%
\end{center}
\vspace{-0.3cm}
\caption{{\bf Bottom}: Examples of changing the background while keeping the caption fixed. {\bf Top}: The respective nearest training images based on pixel-wise L2 distance.}
\label{fig:genimages2}
\vspace{-0.3cm}
\end{figure}

@@

%\vspace{-0.3cm}
\begin{figure}[!h]
\vspace{-0.5cm}
\captionsetup[subfigure]{labelformat=empty}
\begin{center}
\subfloat[\hlcthree{A} rider \hlcone{on} a blue \hlcone{motorcycle} in the \underline{\hlctwo{desert}}.
]{\includegraphics[width=0.23\textwidth]{figures/new/a-rider-on-a-blue-motorcycle-in-the-desert-sharp.png}}\quad
%
\subfloat[\hlcthree{A} rider \hlcone{on} a blue \hlcone{motorcycle} in the \underline{\hlctwo{forest}}.
]{\includegraphics[width=0.23\textwidth]{figures/new/a-rider-on-a-blue-motorcycle-in-the-forest-sharp.png}}\quad
%
\subfloat[\hlctwo{A} \hlcone{surfer}, a woman, and a child walk on the \underline{\hlctwo{beach}}.
]{\includegraphics[width=0.23\textwidth]{figures/new/a-surfer-,-a-woman-,-and-a-child-walk-on-the-beach-sharp.png}}\quad
%
\subfloat[\hlcthree{A} \hlcone{surfer}, a woman, and a child walk on the \underline{\hlczero{sun}}.
]{\includegraphics[width=0.23\textwidth]{figures/new/a-surfer-,-a-woman-,-and-a-child-walk-on-the-sun-sharp.png}}\quad
%
\end{center}
\vspace{-0.3cm}
\caption{Examples of most attended words while changing the background in the caption.}
\label{fig:attentimages1}
\vspace{-0.4cm}
\end{figure}

\vspace{-0.2cm}
\subsection{Comparison With Other Models}
To compare performances of different generative models, we report results on two different metrics as well as a qualitative comparison of different generative models.

\vspace{-0.5cm}
\begin{figure}[!h]
\captionsetup[subfigure]{labelformat=empty}
\begin{center}
\subfloat[alignDRAW]{\includegraphics[width=0.23\textwidth]{figures/new/a-group-of-people-walk-on-a-beach-with-surf-boards-sharp-small.png}}\quad
%
\subfloat[LAPGAN]{\includegraphics[width=0.23\textwidth]{figures/a-group-of-people-walk-on-a-beach-with-surf-boards-lapgan-small.png}}\quad
%
\subfloat[Conv. VAE]{\includegraphics[width=0.23\textwidth]{figures/a-group-of-people-walk-on-a-beach-with-surf-boards-convdeconvvae-small.png}}\quad
%
\subfloat[Fully-conn. VAE]{\includegraphics[width=0.23\textwidth]{figures/a-group-of-people-walk-on-a-beach-with-surf-boards-fcvae-small.png}}\quad
%
\end{center}
\vspace{-0.3cm}
\caption{alignDRAW and three baseline models conditioned on skip-thoughts \cite{skip-thought} displaying results from sampling caption ``A group of people walk on a beach with surf boards". LAPGAN samples look more noisy and it is harder to make out definite objects, whereas the images generated by variational models trained with L2 cost function have a watercolor effect.}
\label{fig:diffmodels}
\vspace{-0.3cm}
\end{figure}

In Figure~\ref{fig:diffmodels}, we generated several samples from the prior of each of the current state-of-the-art generative models, conditioned on the caption ``A group of people walk on a beach with surf boards".

@@

In addition we calculate Structural Similarity Index (SSI), which incorporates luminance and contrast masking into the error calculation. The metric is also calculated on small windows of the image. We sampled 50 images from the prior of each generative model for every test caption and calculated SSI, which is reported in Table~\ref{tab:results}.
\vspace{-0.12in}
\begin{table}[!h]
\small
\begin{center}
\begin{tabulary}{\linewidth}{c || c c c c c || c}
%\hline
%\multicolumn{7}{c}{\textbf{COCO (before sharpening)}} \\
%\hline
& \multicolumn{5}{c||}{Image Search} & Image Similarity \\
%\cline{2-6}
\textbf{Model} & \textbf{R@1} & \textbf{R@5} & \textbf{R@10} & \textbf{R@50} & \textbf{Med r} & \textbf{SSI} \\
\hline
\hline
LAPGAN & - & - & - & - & - & 0.08 \\ % - & - & - & - & - & 0.08
\hline
Fully-conn. VAE (L2 cost) & 1.0 & 5.6 & 10.4 & 51.1 & 51 & 0.156 \\ % 0.688 & 4.58 & 8.74 & 46.832 & 54 & -1
Conv. VAE (L2 cost) & 1.0 & 5.9 & 10.9 & 50.8 & 50 & \textbf{0.164} \\ % 0.672 & 4.448 & 8.476 & 46.596 & 54 & 0.131
skipthoughtDRAW & 2.0 & 11.2 & 18.9 & 63.3 & 36 & 0.157 \\ % 0.988 & 6.18 & 11.48 & 52.512 & 48 & 0.118
noalignDRAW & 2.8 & \textbf{14.1} & \textbf{23.1} & 68.0 & \textbf{31} & 0.155 \\ % 1.0 & 6.124 & 11.18 & 52.112 & 48 & 0.114
alignDRAW & \textbf{3.0} & 14.0 & 22.9 & \textbf{68.5} & \textbf{31} & 0.156 \\ % 0.924 & 6.216 & 11.184 & 52.492 & 48 & 0.115
\end{tabulary}
\end{center}
\vspace{-0.1cm}
\caption{Results of different models on COCO dataset (before sharpening).}
\label{tab:results}
\vspace{-0.3cm}
\end{table}
\vspace{-0.3in}
{\footnotesize
\bibliography{ram-paper}}

>>>>

\maketitle
\vspace*{-0.0cm}
\begin{abstract}
\vspace*{-0.15cm}
Motivated by the recent progress in generative models, we introduce a model that generates images from natural language descriptions. The proposed model iteratively draws patches on a canvas, while attending to the relevant words in the description. After training on MS COCO, we compare our models with several baseline generative models on image generation and retrieval tasks. We demonstrate our model produces higher quality samples than other approaches and generates images with novel scene compositions corresponding to previously unseen captions in the dataset.   
\end{abstract}

@@

Our proposed model 
defines a generative process of images conditioned on the caption. In particular, captions are represented as a sequence of consecutive words and images are represented as a sequence of patches drawn on canvas $c_t$ over time $t=1,...,T$. Our model can be viewed as utilizing the  sequence-to-sequence framework \citep{ilya_mt}.

@@

\vspace{-0.1in}
\subsection{Image Modelling: the Conditional alignDRAW Network}
\vspace{-0.05in}
\comm{
The DRAW network \cite{gregor_draw} is a sequential probabilistic model generating images by accumulating the output at each iterative step. While the original DRAW network assumes the latent variables are independent, it has shown in \citep{bachman_sdm} the model performance is improved by including the dependencies of latent variables. 
}
To generate images conditioned on the caption information, we extended the DRAW network \citep{gregor_draw} to include caption representation $\hlangall$ at each step, shown in Figure~\ref{fig:figmodel}. The conditional DRAW network is a stochastic recurrent neural network that consists of a set of latent variables $\Lat_t$ at each time step. Unlike the original DRAW network where latent variables are independent unit Gaussians $\mathcal{N}(0, I)$, the latent variables in the proposed alignDRAW model have their mean and variance depend on the previous recurrent hidden states $\hdec_{t-1}$ as in \cite{bachman_sdm}. Formally, the image is generated by 
iteratively computing the following equations for $t=1,...,T$ (see Figure~\ref{fig:figmodel}):
\begin{align}
\vspace{-0.1in}
\label{eq:x_hat}
\lat_t &\sim \prior(\Lat_t|\Lat_{1:t-1}) = \mathcal{N}(\mu_{t}(\hdec_{t-1}), \sigma_{t}(\hdec_{t-1})),\\
\hdec_t &= \decoder(\hdec_{t-1}, z_t, s_{t-1}), \\
s_{t} = align(\hdec_{t-1}, & \hlangall); \quad \quad
\canv_t = \canv_{t-1} + \writeop(\hdec_t), 
\label{eq:write}
%\canv_t &= \canv_{t-1} + \writeop(\hdec_t)
\end{align}
where $\writeop$ and $\readop$ are the same attention operators as in \citep{gregor_draw}.
The $align$ function is used to compute the alignment between the input caption and intermediate image generative steps \citep{bahdanau_mt}. Given the caption representation from the language model, $\hlangall = [\hlang_{1}, \hlang_{2}, ..., \hlang_{N}]$, the $align$ operator outputs a dynamic sentence representation $s_t$ at each step through a weighted sum using alignment probabilities $\alpha_{1...N}$:
%\begin{align}
%\vspace{-0.1in}
$s_t=align(\hdec_{t-1}, \hlangall) = \alpha_{1}\hlang_{1} + \alpha_{2}\hlang_{2} + ... + \alpha_{N}\hlang_{N}.$
%\end{align}
The corresponding alignment probabilities $\alpha_{1...N}$ at each step are obtained using the caption representation $\hlangall$ and the hidden state of the generative model $\hdec_{t}$, as in \citep{bahdanau_mt}.
\comm{
\begin{align}
e_{tj} &= v^{T}tanh(U\hlang_{j} + W\hdec_{t} + b)\\
\alpha_{j} &= \frac{exp(e_{tj})}{\sum_{j=1}^{N}exp(e_{tj})}.
\end{align}
}
\comm{
Here $\hlang_{0}$ is initialized to the learned bias.
Setting $\alpha_{1...N}$ to $\frac{1}{N}$ turns the encoder into the vanilla model introduced in \citep{cho_mt} without the attention. 
}

\vspace{-0.05cm}
\subsection{Learning and Generation}
\vspace{-0.05in}

\begin{figure}[!h]
\vspace{-0.45in}
\captionsetup[subfigure]{labelformat=empty}
\begin{center}
\includegraphics[width=0.8\textwidth]{figures/alignDrawAnnotated.pdf}\quad
%
\end{center}
\vspace{-0.2in}
\caption{alignDRAW: Generative model of images conditioned on captions.}
\label{fig:figmodel}
\vspace{-0.2in}
\end{figure}
%\vspace{-0.1in}

Our conditional alignDRAW model is trained to maximize the variational lower bound on the log-likelihood, $\log \sum_{\Lat_{1:T}} \prior(\Lat_{1:T})p(\oimage\given\icaption, \Lat_{1:T})$. The posterior inference is approximated by an inference RNN $\post(\Lat_{1:T}\given\icaption,\oimage)$ shown in the red dashed box in Figure~\ref{fig:figmodel}.   
Overall, the variational objective $\loss$ is defined with the model parameters vector $\theta$ as follows:
\begin{align}
\vspace{-0.1in}
\small
\loss_\theta =  &\expectation_{\post(\Lat_{1:t}\given\icaption,\oimage)}\left[ - \log p(\oimage\given\icaption,\Lat_{1:T}) + \sum_{t=2}^T\kldiv\left(\post(\Lat_t\given\Lat_{1:t-1},\icaption,\oimage)\klBars\prior(\Lat_t\given\Lat_{1:t-1},\icaption)\right) \right] \nonumber \\& + \kldiv\left(\post(\Lat_1\given\oimage)\klBars\prior(\Lat_1\given\icaption)\right).
\end{align}

@@

\vspace{-0.2cm}
\section{Experiments on MS COCO dataset}
\vspace{-0.1in}
In the following subsections, we analyze both the qualitative and quantitative aspects of our model as well as compare its performance with that of other, related generative models\footnote{To see more generated images, go to \url{http://www.cs.toronto.edu/~emansim/cap2im.html}}.
First, we wanted to see whether the model understood one of the most basic properties of any object, the color. In Figure~\ref{fig:genimages1}, we generated images of school buses with four different colors: yellow, red, green and blue. Although, there are images of buses with different colors in the training set, all mentioned school buses are specifically colored yellow. Despite that, the model managed to generate images of an object that is visually reminiscent of a school bus that is painted with the specified color.

% \vspace{-0.5cm}
\begin{figure}[!h]
 \vspace{-0.4cm}
\captionsetup[subfigure]{labelformat=empty}
\begin{center}
\subfloat[A \underline{yellow} school bus parked in a parking lot.
]{\includegraphics[width=0.23\textwidth]{figures/new/a-yellow-school-bus-parked-in-a-parking-lot-sharp.png}}\quad
%
\subfloat[A \underline{red} school bus parked in a parking lot.
]{\includegraphics[width=0.23\textwidth]{figures/new/a-red-school-bus-parked-in-a-parking-lot-sharp.png}}\quad
%
\subfloat[A \underline{green} school bus parked in a parking lot.
]{\includegraphics[width=0.23\textwidth]{figures/new/a-green-school-bus-parked-in-a-parking-lot-sharp.png}}\quad
%
\subfloat[A \underline{blue} school bus parked in a parking lot.
]{\includegraphics[width=0.23\textwidth]{figures/new/a-blue-school-bus-parked-in-a-parking-lot-sharp.png}}\quad
%
\end{center}
\vspace{-0.2cm}
\caption{Examples of changing the color while keeping the caption fixed. Best viewed in colour.}
\label{fig:genimages1}
\vspace{-0.3cm}
\end{figure}
Apart from changing the colors of objects, we experimented with changing the background of the scene described in a caption to see whether this would result in the appropriate changes in the generated samples. The task of changing the background of an image is somewhat harder than just changing the color of an object because the model will have to make alterations over a wider visual area. Nevertheless, as shown in Figure~\ref{fig:genimages2}, changing the skies from blue to rainy in a caption as well as changing the grass type from dry to green in another caption resulted in the appropriate changes in the generated image. The nearest images from the training set also indicate that the model was not simply copying the patterns it observed during the learning phase.

\vspace{-0.4cm}
\begin{figure}[!h]
\captionsetup[subfigure]{labelformat=empty}
\begin{center}
\subfloat[]{\includegraphics[width=0.23\textwidth]{figures/new/a-very-large-commercial-plane-flying-in-blue-skies-closest.png}}\quad
%
\subfloat[]{\includegraphics[width=0.23\textwidth]{figures/new/a-very-large-commercial-plane-flying-in-rainy-skies-closest.png}}\quad
%
\subfloat[]{\includegraphics[width=0.23\textwidth]{figures/new/a-herd-of-elephants-walking-across-a-dry-grass-field-closest.png}}\quad
%
\subfloat[]{\includegraphics[width=0.23\textwidth]{figures/new/a-herd-of-elephants-walking-across-a-green-grass-field-closest.png}}\\
\vspace{-0.45cm}
%
\subfloat[A very large commercial plane flying in \underline{blue} skies.
]{\includegraphics[width=0.23\textwidth]{figures/new/a-very-large-commercial-plane-flying-in-blue-skies-sharp.png}}\quad
%
\subfloat[A very large commercial plane flying in \underline{rainy} skies.
]{\includegraphics[width=0.23\textwidth]{figures/new/a-very-large-commercial-plane-flying-in-rainy-skies-sharp.png}}\quad
%
\subfloat[A herd of elephants walking across a \underline{dry} grass field.
]{\includegraphics[width=0.23\textwidth]{figures/new/a-herd-of-elephants-walking-across-a-dry-grass-field-sharp.png}}\quad
%
\subfloat[A herd of elephants walking across a \underline{green} grass field.
]{\includegraphics[width=0.23\textwidth]{figures/new/a-herd-of-elephants-walking-across-a-green-grass-field-sharp.png}}\quad
%
\end{center}
\vspace{-0.3cm}
\caption{{\bf Bottom}: Examples of changing the background while keeping the caption fixed. {\bf Top}: The respective nearest training images based on pixel-wise L2 distance. Best viewed in colour.}
\label{fig:genimages2}
\vspace{-0.3cm}
\end{figure}

@@

%\vspace{-0.3cm}
\begin{figure}[!h]
\vspace{-0.5cm}
\captionsetup[subfigure]{labelformat=empty}
\begin{center}
\subfloat[\hlcthree{A} rider \hlcone{on} a blue \hlcone{motorcycle} in the \underline{\hlctwo{desert}}.
]{\includegraphics[width=0.23\textwidth]{figures/new/a-rider-on-a-blue-motorcycle-in-the-desert-sharp.png}}\quad
%
\subfloat[\hlcthree{A} rider \hlcone{on} a blue \hlcone{motorcycle} in the \underline{\hlctwo{forest}}.
]{\includegraphics[width=0.23\textwidth]{figures/new/a-rider-on-a-blue-motorcycle-in-the-forest-sharp.png}}\quad
%
\subfloat[\hlctwo{A} \hlcone{surfer}, a woman, and a child walk on the \underline{\hlctwo{beach}}.
]{\includegraphics[width=0.23\textwidth]{figures/new/a-surfer-,-a-woman-,-and-a-child-walk-on-the-beach-sharp.png}}\quad
%
\subfloat[\hlcthree{A} \hlcone{surfer}, a woman, and a child walk on the \underline{\hlczero{sun}}.
]{\includegraphics[width=0.23\textwidth]{figures/new/a-surfer-,-a-woman-,-and-a-child-walk-on-the-sun-sharp.png}}\quad
%
\end{center}
\vspace{-0.2cm}
\caption{Examples of most attended words while changing the background in the caption.}
\label{fig:attentimages1}
\vspace{-0.3cm}
\end{figure}

\vspace{-0.2cm}
\subsection{Comparison With Other Models}
\vspace{-0.1in}
To compare performances of different generative models, we report results on two different metrics as well as a qualitative comparison of different generative models.
In Figure~\ref{fig:diffmodels}, we generated several samples from the prior of each of the current state-of-the-art generative models, conditioned on the caption ``A group of people walk on a beach with surf boards".

\vspace{-0.3cm}
\begin{figure}[!h]
\captionsetup[subfigure]{labelformat=empty}
\begin{center}
\subfloat[alignDRAW]{\includegraphics[width=0.23\textwidth]{figures/new/a-group-of-people-walk-on-a-beach-with-surf-boards-sharp-small.png}}\quad
%
\subfloat[LAPGAN]{\includegraphics[width=0.23\textwidth]{figures/a-group-of-people-walk-on-a-beach-with-surf-boards-lapgan-small.png}}\quad
%
\subfloat[Conv. VAE]{\includegraphics[width=0.23\textwidth]{figures/a-group-of-people-walk-on-a-beach-with-surf-boards-convdeconvvae-small.png}}\quad
%
\subfloat[Fully-conn. VAE]{\includegraphics[width=0.23\textwidth]{figures/a-group-of-people-walk-on-a-beach-with-surf-boards-fcvae-small.png}}\quad
%
\end{center}
\vspace{-0.2cm}
\caption{alignDRAW and three baseline models conditioned on skip-thoughts \cite{skip-thought} displaying results from sampling caption ``A group of people walk on a beach with surf boards". LAPGAN samples look more noisy and it is harder to make out definite objects, whereas the images generated by variational models trained with L2 cost function have a watercolor effect.}
\label{fig:diffmodels}
\vspace{-0.2cm}
\end{figure}

%In Figure~\ref{fig:diffmodels}, we generated several samples from the prior of each of the current state-of-the-art generative models, conditioned on the caption ``A group of people walk on a beach with surf boards".

@@

In addition we calculate Structural Similarity Index (SSI), which incorporates luminance and contrast masking into the error calculation. The metric is also calculated on small windows of the image. We sampled 50 images from the prior of each generative model for every test caption and calculated SSI, which is reported in Table~\ref{tab:results}.
\vspace{-0.1in}
\begin{table}[!h]
\small
\begin{center}
\begin{tabulary}{\linewidth}{c || c c c c c || c}
%\hline
%\multicolumn{7}{c}{\textbf{COCO (before sharpening)}} \\
%\hline
& \multicolumn{5}{c||}{Image Search} & Image Similarity \\
%\cline{2-6}
\textbf{Model} & \textbf{R@1} & \textbf{R@5} & \textbf{R@10} & \textbf{R@50} & \textbf{Med r} & \textbf{SSI} \\
\hline
\hline
LAPGAN & - & - & - & - & - & 0.08 \\ % - & - & - & - & - & 0.08
\hline
Fully-conn. VAE (L2 cost) & 1.0 & 5.6 & 10.4 & 51.1 & 51 & 0.156 \\ % 0.688 & 4.58 & 8.74 & 46.832 & 54 & -1
Conv. VAE (L2 cost) & 1.0 & 5.9 & 10.9 & 50.8 & 50 & \textbf{0.164} \\ % 0.672 & 4.448 & 8.476 & 46.596 & 54 & 0.131
skipthoughtDRAW & 2.0 & 11.2 & 18.9 & 63.3 & 36 & 0.157 \\ % 0.988 & 6.18 & 11.48 & 52.512 & 48 & 0.118
noalignDRAW & 2.8 & \textbf{14.1} & \textbf{23.1} & 68.0 & \textbf{31} & 0.155 \\ % 1.0 & 6.124 & 11.18 & 52.112 & 48 & 0.114
alignDRAW & \textbf{3.0} & 14.0 & 22.9 & \textbf{68.5} & \textbf{31} & 0.156 \\ % 0.924 & 6.216 & 11.184 & 52.492 & 48 & 0.115
\end{tabulary}
\end{center}
\vspace{-0.1cm}
\caption{Results of different models on COCO dataset (before sharpening).}
\label{tab:results}
\vspace{-0.3cm}
\end{table}
\vspace{-0.3in}
{\footnotesize
\bibliography{ram-paper}}
